---
title: "ExData_Project2"
author: "Garrett Farley"
date: "5/21/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(readr)
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/exdata%2Fdata%2FNEI_data.zip",temp)
NEI<-read_rds("summarySCC_PM25.rds")
SCC<-read_rds("Source_Classification_Code.rds")
unlink(temp)
```

## Analysis Questions

### Have total emissions from PM2.5 decreased in the United States from 1999 to 2008? Using the base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

```{r echo = FALSE}
#Sum total emissions by year
total_em_by_year<-NEI%>%group_by(year)%>%summarise(sums = sum(Emissions))

#Plot total emissions
with(total_em_by_year, 
     plot(year, sums, type = "l", 
          main = "Total Emissions by Year", 
          xlab = "Year", 
          ylab = "Total Emissions"))
```

### Have total emissions from PM2.5 decreased in the Baltimore City, Maryland from 1999 to 2008?

```{r echo = FALSE}
#Filter and summarise Baltimore emissions data
baltimore<-NEI%>%filter( fips == "24510")%>%
        group_by(year)%>%
        summarise(sums = sum(Emissions))

#Plot data
with(baltimore, plot(year, sums, type = "l", 
                     main = "Total Emissions in Baltimore by Year",
                     xlab = "Year", 
                     ylab = "Total Emissions"))
```

### Of the four types of sources indicated by the point, nonpoint, onroad, nonroad variable, which of these four sources have seen decreases in emissions from 1999–2008 for Baltimore City? Which have seen increases in emissions from 1999–2008?

```{r echo = FALSE}
#Filter and summarise Baltimore emissions by type
baltimore_by_type<-NEI%>%filter(fips == "24510")%>%
        group_by(type, year)%>%
        summarise(sums = sum(Emissions))

#Plot data
b<-ggplot(data = baltimore_by_type, aes(x = year, y = sums, group = type, color = type))+
        geom_line()+
        labs(x = "Year",
             y = "Total Emissions",
             title = "Total Emissions by Type by Year in Baltimore")
b
```

### Across the United States, how have emissions from coal combustion-related sources changed from 1999–2008?

```{r echo = FALSE}
#Use coal related SCC to subset NEI and summarise by group
coal_comb_source<-SCC%>%filter(grepl(pattern = "Coal", x = EI.Sector))
coal_scc<-coal_comb_source%>%distinct(SCC)
coal<- subset(NEI, coal_scc$SCC %in% NEI$SCC)
coal_by_year<-coal%>%group_by(year)%>%
        summarise(sums = sum(Emissions))
#Plot data
c<-ggplot(data = coal_by_year, aes(x = year, y = sums))+
        geom_line()+
        labs(x = "Year", 
             y = "Total Emissions",
             title = "Coal Emissions by Year")
c

```

### How have emissions from motor vehicle sources changed from 1999–2008 in Baltimore City?

```{r echo=FALSE}
#Use motor vehicle related SCC to subset NEI and then summarise by group
motor_vehicle_source<-SCC%>%filter(grepl(pattern = "Mobile Sources", x = SCC.Level.One))
motor_scc<-motor_vehicle_source%>%distinct(SCC)
motor<-subset(NEI, motor_scc$SCC %in% NEI$SCC)
motor_by_year_balt<-motor%>%
        filter(fips == "24510")%>%
        group_by(year)%>%
        summarise(sums = sum(Emissions))

#Plot data
m<-ggplot(data = motor_by_year_balt, aes(x = year, y = sums))+
                 geom_line()+
        labs(x = "Year", 
             y = "Total Emissions", 
             title = "Total Motor Vehicle Emissions in Baltimore by Year")
m
```

### Compare emissions from motor vehicle sources in Baltimore City with emissions from motor vehicle sources in Los Angeles County, California (fips == "06037"). Which city has seen greater changes over time in motor vehicle emissions?


```{r echo = FALSE}
#Use previous motor subset and filter for Baltimore and Los Angeles fips
motor_by_year_los_balt<-motor%>%
        filter(fips == "24510" | fips == "06037")%>%
        group_by(fips, year)%>%
        summarise(sums = sum(Emissions))

#Plot data
mlb<-ggplot(data = motor_by_year_los_balt, aes(x = year, y = sums, group = fips, color = fips))+
                          geom_line()+
        labs(x = "Year", 
             y = "Total Emissions", 
             title = "Vehicle Emissions for LA and Baltimore by Year")
mlb
```
